The Analog Method for Precipitation Downscaling: Finding Better Analog Situations at a Sub-Daily Time Step

### 2016-11-10 Moving temporal window - HESS/template.tex
HESS paper under corrections.

The first method is based on the analogy of the atmospheric circulation only \citep[Table \ref{table:method_2Z},][]{Obled2002, Bontron2005, Marty2012}. Searching analog situations for a target day starts by a preselection step of the potential candidate for analogy. Here, it has been limited to the 4 months centered around the target date for every year of the archive, in order to cope with seasonal effects. Then, the similarity of the atmospheric circulation of the target date with every day of the preselection is assessed by processing the \citet{Teweles1954} score (S1), which is a comparison of gradients, over a certain spatial window and at certain hours. The smaller the S1 values, the more similar the pressure fields.

@@

\begin{equation}
\label{eq:CRPSS}
S_{\text{SCRP}} = \frac{S_{\text{CRP}}-S_{\text{CRP}}^{r}}{S_{\text{CRP}}^{p}-S_{\text{CRP}}^{r}} = 1-\frac{S_{\text{CRP}}}{S_{\text{CRP}}^{r}}
\end{equation}
where $S_{\text{CRP}}^{r}$ is the $S_{\text{CRP}}$ value for the reference (climatological distribution) and $S_{\text{CRP}}^{p}$ would be the one for a perfect prediction (which implies $S_{\text{CRP}}^{p}~=~0$). A better prediction is characterized by an increase in CRPSS.

@@

When searching for analogs on the geopotential heights, as in the 2Z method, there are now 4 times more candidates than before, which obviously allows to find better matches.

Figure \ref{fig:changes_S1_analogs} presents the changes in the distributions of the S1 criterion for the $1^{st}$, $5^{th}$, $20^{th}$ and $40^{th}$ analogs for the Ulrichen station on the whole calibration period, due to introduction of the MTW. The precipitation target remains as before, that is centered on 18~h~UTC (6~h~UTC to 6~h~UTC the next day). The shapes of the distributions of the conventional approach and the MTW are similar, but the values of the analogy criteria are now reduced (shifted to the left), and therefore better. An increase in the difference between a fixed window and a moving window is identifiable, which means that the last analogs are further improved. The latter effect is due to the accumulation of improvements brought by the new analog situations in the selection.

@@

When considering the second level of analogy of the 2Z-2MI method (Table \ref{table:method_2Z-2MI}), the number of candidates situations did not increase, as they are subsampled in the previously $N_{1}$ selected analogs, but their dates have changed. In contrast to earlier, both a reduction or an increase of the RMSE analogy criterion values are possible. Indeed, Fig. \ref{fig:changes_RMSE} shows an almost insignificant improvement of the RMSE values. Unlike for the first level of analogy, the relative improvement of the RMSE values are distributed relatively symmetrically around zero, with improvements and losses of the same amplitude. Once again, the results for the other stations are similar.

@@

A systematic improvement of the S1 values was previously observed.

The new performance scores are provided in Fig. \ref{fig:plots_CRPSS_2Z} and \ref{fig:plots_CRPSS_2Z-2MI}, along with the differences regarding the conventional method with the same archive length. The differences ranges from 0.57 to 2.14 points for the 2Z method and from 1.53 to 2.20 points for the 2Z-2MI method. The introduction of the MTW brings an improvement of the performance that is not very large, but that is nevertheless significant. Moreover, it requires no additional predictor. No relationship was found between the improvement of the score and the reduction of the S1 criteria, neither with the season.

@@

The increasing positive trend of skill improvements regarding the precipitation threshold is significant and shows that the prediction of higher precipitation totals is further improved. Thus, both the analogy criteria and the performance scores are improved to a greater extent for heavier precipitation events. On the contrary, the non-rainy days and small accumulations are not improved.

@@

After recalibrating, some changes in the optimal number of analogs can be observed for both 2Z and 2Z-2MI methods (Table \ref{table:analog_nb}). Among these, the east-west dimension of the spatial windows of the circulation analogy tends to decrease (not shown). More significantly, the optimal number of analogs increases after introducing the MTW, of a significant order of magnitude: 25~\% to 83~\% for the 2Z method and 20~\% 67~\% for the 2Z-2MI method. The number of analogs of the first analogy level of the 2Z-2MI method even reached three times its previous value for the Visp station.

This increase in number of analogs has a slight effect on the performance of the different precipitation classes. The same analysis as in Sect. \ref{sec:improvement_CRPSS_precip_threshold} has been performed again on the basis of the newly calibrated parameters. The results are generally very similar, but a slight performance increase of small precipitation values can be observed at the expense of higher amounts. The change in the analogs numbers is likely to be responsible for this difference in the performance distributions.

The values of the CRPSS scores for both methods (Fig. \ref{fig:plots_CRPSS_2Z} and \ref{fig:plots_CRPSS_2Z-2MI}) have significantly increased after recalibration.

@@

@@

It can be assumed that the atmospheric conditions with a low dynamism, such as the frequent anticyclonic situations, will not be radically improved by the introduction of the MTW. Conversely, dynamic situations, such as weather disturbances, have a well marked temporal evolution. Indeed, the position of the driving elements such as the low-pressure center and the fronts change significantly during a day. We can therefore expect to improve more significantly these situations with a higher dynamism when introducing a MTW, as better matches to the target situation may be found.

The dynamism of a given atmospheric condition cannot be easily quantified. A basic assumption is considered here, which stipulates that the more a day is rainy, the more dynamic the situation is.

@@

\subsection{Increase in number of analog situations}

It seems as if the method perceive the increase in the archive length and diversity by allowing to extract more analogs. As we saw in Fig. \ref{fig:changes_S1}, the improvement of the S1 criteria grows along with the rank of the analog, which shows an accumulation of better analog situations in the distributions. It seems thus profitable to widen the selection of analogs in order to keep also some whose rank has increased, as they appear to be also relevant to predict precipitation values. The number of good analogs is thus globally increased.

\subsection{Why not just making a 6-hourly prediction ?}

One can question the interest of using daily totals when a 6-hourly precipitation series can be used. The first reason is that the 6-hourly time series generated by the AM may not represent well the dynamic of the accurate precipitation (results not shown), due to a smoothed signal. Then, sometimes one just does not need finer resolution than the daily time step. Finally, when using a reconstructed precipitation archive, the errors in intra-day precipitation distributions have a lesser impact over 24~h daily totals.

@@

The considered reanalysis dataset, which is the NCEP/NCAR reanalysis I \citep{Kalnay1996}, is relatively old and has a coarse resolution. Newer reanalysis dataset could have been used to achieve this study, and might have resulted in higher performance scores. However, one can assume that the gain observed here by introducing a MTW should also be found for a better reanalysis dataset, especially if the information of the atmospheric circulation is more accurate.

>>>>

The first method is based on the analogy of the atmospheric circulation only \citep[Table \ref{table:method_2Z},][]{Obled2002, Bontron2005, Marty2012}. Searching analog situations for a target day starts by a preselection step of the potential candidate for analogy. Here, it has been limited to the 4 months centered around the target date for every year of the archive, in order to cope with seasonal effects (seasonal stratification). Then, the similarity of the atmospheric circulation of the target date with every day of the preselection is assessed by processing the \citet{Teweles1954} score (S1), which is a comparison of gradients, over a certain spatial window and at certain hours. The smaller the S1 values, the more similar the pressure fields.

@@

\begin{equation}
\label{eq:CRPSS}
S_{\text{SCRP}} = \frac{S_{\text{CRP}}-S_{\text{CRP}}^{r}}{S_{\text{CRP}}^{p}-S_{\text{CRP}}^{r}} = 1-\frac{S_{\text{CRP}}}{S_{\text{CRP}}^{r}}
\end{equation}
where $S_{\text{CRP}}^{r}$ is the $S_{\text{CRP}}$ value for the reference (climatological distribution) and $S_{\text{CRP}}^{p}$ would be the one for a perfect prediction (which implies $S_{\text{CRP}}^{p}~=~0$). A better prediction is characterized by an increase in CRPSS: $S_{\text{SCRP}}~=~1$ for a perfect prediction, and $S_{\text{SCRP}}~<~0$ for a prediction with a lower skill than the reference.

@@

When searching for analogs on the geopotential heights, as in the 2Z method, there are now 4 times more candidates (even though not fully independent) than before, which obviously allows to find better matches.

Figure \ref{fig:changes_S1_analogs} presents the changes in the distributions of the S1 criterion for the $1^{st}$, $5^{th}$, $20^{th}$ and $40^{th}$ analog situations for the Ulrichen station on the whole calibration period (1961-2008), due to introduction of the MTW. The precipitation target remains as before, that is centered on 18~h~UTC (6~h~UTC to 6~h~UTC the next day). The shapes of the distributions of the conventional approach and the MTW are similar, but the values of the analogy criteria are now reduced (shifted to the left), and therefore better. For analogs with higher ranks (e.g. $20^{th}$ or $40^{th}$), the difference between the two distributions is larger than for the first ranks, which means that the improvement increases with the rank of the analogs. This is due to the accumulation of improvements brought by the new analog situations in the selection.

@@

When considering the second level of analogy of the 2Z-2MI method (Table \ref{table:method_2Z-2MI}), the number of candidates situations did not increase, as they are subsampled in the previously $N_{1}$ selected analogs, but their dates have changed. In contrast to the AM on the atmospheric circulation only, both a reduction or an increase of the RMSE analogy criterion values are possible here with the MTW compared to the static approach. Indeed, Fig. \ref{fig:changes_RMSE} shows an almost insignificant improvement of the RMSE values. Unlike for the first level of analogy, the relative improvement of the RMSE values are distributed relatively symmetrically around zero, with improvements and losses of the same amplitude. Once again, the results for the other stations are similar.

@@

A systematic improvement of the analogy criteria values was previously observed.

The new performance scores are provided in Fig. \ref{fig:plots_CRPSS_2Z} and \ref{fig:plots_CRPSS_2Z-2MI} for the 2Z and 2Z-2MI methods respectively. The gains, relatively to the static approach on the reduced archive, range from 0.57 to 2.14 points for the 2Z method and from 1.53 to 2.20 points for the 2Z-2MI method.

@@

The performance score is improved for days with higher precipitation after the introduction of the MTW. On the contrary, non-rainy days and small accumulations are not improved.

@@

After recalibrating, changes in the AM parameters can be observed for both 2Z and 2Z-2MI methods. Among these, the zonal extent of the spatial windows of the circulation analogy tends to decrease (not shown). More significantly, the optimal number of analogs increases after introducing the MTW, of a significant order of magnitude: 25~\% to 83~\% for the 2Z method and 20~\% 67~\% for the 2Z-2MI method (Table \ref{table:analog_nb}). The number of analogs of the first analogy level of the 2Z-2MI method even reached three times its previous value for the Visp station.

The values of the CRPSS scores for both methods (Fig. \ref{fig:plots_CRPSS_2Z} and \ref{fig:plots_CRPSS_2Z-2MI}) have significantly increased after recalibration. When analyzing the change in performance per precipitation classes, the results (not shown) are very similar to what was observed in Sect. \ref{sec:improvement_CRPSS_precip_threshold}, with a slight performance increase of small precipitation values can be observed at the expense of higher amounts.

@@

The introduction of the MTW brings an improvement of the performance that is not very large, but that is nevertheless significant. Moreover, it requires no additional predictor. More complex methods with additional levels of analogy and more predictors do increase the performance score, and the MTW could also be applied in these situations as it is likely to improve the analogy of the atmospheric circulation anyway.

No relationship was found between the improvement of the score and the reduction of the S1 criteria, neither with the season.

@@

Both the analogy criteria (Sect. \ref{sec:influence_precip}) and the performance scores (Sect. \ref{sec:improvement_CRPSS_precip_threshold}) are improved to a greater extent for heavier precipitation events. This is likely due to the fact that higher precipitation events are a consequence of atmospheric conditions with a higher dynamism, such as weather disturbances, which have a well marked temporal evolution. Indeed, the position of the driving elements such as the low-pressure center and the fronts change significantly during a day. We can therefore expect to improve more significantly these situations with a higher dynamism when introducing a MTW, as better matches to the target situation may be found. On the opposite, days with a low dynamism of the atmospheric circulation, such as anticyclonic situations, will not be radically improved by the introduction of the MTW.

@@

\subsection{On the increase in number of analogs}

After the recalibration of the method with the MTW, the optimal analog numbers were significantly higher than for the original method.

As we saw in Fig. \ref{fig:changes_S1}, the improvement of the S1 criteria grows along with the rank of the analog, which shows an accumulation of better analog situations in the distributions. It seems thus profitable to widen the selection of analogs in order to keep also some whose rank has increased, as they appear to be also relevant to predict precipitation values. The number of good analogs is thus globally increased.

The change in the analogs numbers is likely to be responsible for this difference in the performance distributions. --> thresholds

\subsection{Why not just making a 6-hourly predictions ?}

One can question the interest of using moving daily totals when 6-hourly precipitation series can be predicted instead. A first reason is that the 6-hourly time series generated by the AM might not represent accurately the intra-daily precipitation distribution (results not shown). Or, sometimes one just does not need finer resolution than the daily time step, eventually using another disaggregation technique afterwards. Finally, when using a reconstructed precipitation archive, the errors in intra-day precipitation distributions have a lesser impact over 24~h daily totals.

@@

The considered reanalysis dataset, which is the NCEP/NCAR reanalysis I \citep{Kalnay1996}, is relatively old and has a coarse resolution. Newer reanalysis dataset could have been used to achieve this study, and might have resulted in higher performance scores. However, one can assume that the gain observed here by introducing a MTW should also be found for a better reanalysis dataset, especially if the information of the atmospheric circulation is more accurate. Indeed, a better dataset does not negatively influence the fact that one can find better analog situations at other hours of the day rather than at fixed hours. A comparative analysis of several reanalysis dataset within the AM is under work and will be the topic of a dedicated paper.
